Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2022-08-11 manuscript/Chapter3-manuscript.Rmd
move dump ideas to tests. update intro MS.

The integration of individual behavioral variation into studies of predator-prey interactions has been championned for some time, and a series of empirical papers have revelealed important insights on its consequences for functional responses, habitat use, and prey choice. However, a recurring challenge impeding empirical studies of predator-prey behavioral interactions is the difficulty of collecting data simultaneously on both the predator and prey. We have recently demonstrated that using an online predator-prey videogame can help uncover subtle details about the mechanisms shaping these behavioral interactions [@FraserFranco].

*ici, mettre le paragraphe sur la variation individuelle, et ensuite expliquer ce qu'on cherche à comprendre dans la littérature écologique comportementale*

@@

*un autre truc que je pourrais parler est le fait que si on assume que les individus deviennent meilleurs avec l'expertise (Dukas), on voit pourtant dans les études marines que les individus de groupes d'âges variés performent souvent de manière similaire. Ça amènerait bien l'autre paragraphe pour montrer que les interactions écologiques ont un rôle clé!!*

*ou juste dire que pour comprendre le role de l'expertise, il faut s'attarder à l'étudier avec la specialisation et voir les consequences pour la fitness ou issue des interactions predateur proie. ca introduit mieux je crois. dire genre que une théorie prédit que spécialisation est avantage quand proies peu variable, et ensuite ça lie bien l'hypothèse de variabilité que je décris en bas*

A number of long-term studies observed that generalist and specialist individuals can achieve similar fitness [sources: à partir de Woo]. An emerging explanation is that temporal fluctuations in the predictability of resources (abundance, availability, behaviour) may favor one or the other strategy depending on time scales [sources : Woo, Phillips, Ceia, etc, Courbin, Chang]. The resource-predictability hypothesis advances that when resources are predictable, particularly on short time-scales, individual specialists should benefit from higher fitness returns. The rationale is that individuals should have higher prey delivery rates when they repeatedly employ the same foraging technique, as it is easier to assess prey predictability over shorter time-scales. In contrast, individual generalists should be advantaged over longer time scales, as resource parameters are expected to fluctuate. [revoir ceia et ramos pour les sources]
*ce qu'il manque, c'est qu'on a aucune idée comment des différences individuelles de variabilité environnementale se traduisent par des différences de stratégies*

@@

 *Difficile de mesurer à travers le temps blabla. J'introduis ensuite le jeu vidéo* *vidfeogames offer a good alternative to lab studies because proche de wild = observational*
*(idéee: par contre, décomposer l'âge de l'expérience est difficile, et particulièrement chez des espèces sur le terrain puisque ça peut impliquer des techniques invasives)*

Here, we collected individual behavioural data from players in *Dead by Daylight* to investigate how hunting experience and the behavioural variation in prey encountered shaped predator foraging specialization. This enabled us to have a high degree of precision on the interaction as both the behaviour of the predator and the four prey were measured simultaneously in each trial. First, we hypothesize that individual predator behaviour will change with hunting experience. If experience reduces the costs of switching between hunting tactics, then predators should become more flexible through time. Alternatively, if experience enables the refinement of the hunting tactics, then individuals may instead specialize through time. Second, we hypothesize that prey behaviour will shape the developmental trajectories of the predator hunting tactics. Whether predators specialize or not with experience may depend on the behaviour of their prey, for instance, as it may be harder to specialize when encounters with prey are more variable. Third, if individuals differ in their degree of foraging specialization, then we expect that the success of foraging specialists and generalists will be equal. Specialist hunters should fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.

>>>>

The integration of individual behavioral variation into studies of predator-prey interactions has gained traction in recent years, with a series of empirical papers revealing important consequences for functional responses, habitat use, and prey choice (REFS). However, an important and recurring challenge impeding empirical studies of predator-prey behavioral interactions, at the individual level, is the need of collecting data simultaneously on both the predator and prey. We have recently demonstrated with behavioural data from an online predator-prey videogame that virtual systems can uncover subtle details on the mechanisms shaping these interactions [@FraserFranco.etal2022]. For instance, we tested the classical locomotor crossover hypothesis championed by @Huey.Pianka1981, and found support of its predictions only for active hunters, which concords with a recent experiment involving assassin bugs [@Matsumura.Miyatake2022a]. Thus, general ecological patterns can potentially be derived from these virtual systems and may help in advacing the current gaps in predator-prey research. Other examples of virtual ecological studies show that predation regimes can drive individual variation in risk perception [@Beauchamp2020], that familiarity with prey partners has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes [Santostefano et al. in prep]. Because these systems generate large volumes of data on prey and predators simultaneously throughout their lifetime in the game, they offer the opportunity of uncovering how expertise and prey behavior shape individual predator foraging specialization and its potential fitness consequences.

@@

*ou juste dire que pour comprendre le role de l'expertise, il faut s'attarder à l'étudier avec la specialisation et voir les consequences pour la fitness ou issue des interactions predateur proie. ca introduit mieux je crois. dire genre que une théorie prédit que spécialisation est avantage quand proies peu variable, et ensuite ça lie bien l'hypothèse de variabilité que je décris en bas. faire ressortir qu'il n'y a pas de consencus. There is right now no concencus as to whether blabla, (examples)*

Some long-term studies show that generalist and specialist individuals can achieve similar fitness [sources: à partir de Woo]. An emerging explanation is that temporal fluctuations in the predictability of resources (abundance, availability, behaviour) may favor one or the other strategy depending on time scales [sources : Woo, Phillips, Ceia, etc, Courbin, Chang]. The resource-predictability hypothesis advances that when resources are predictable, particularly on short time-scales, individual specialists should benefit from higher fitness returns. The rationale is that individuals should have higher prey delivery rates when they repeatedly employ the same foraging technique, as it is easier to assess prey predictability over shorter time-scales. In contrast, individual generalists should be advantaged over longer time scales, as resource parameters are expected to fluctuate. [revoir ceia et ramos pour les sources]
*ce qu'il manque, c'est qu'on a aucune idée comment des différences individuelles de variabilité environnementale se traduisent par des différences de stratégies*

@@

Here, we analyze individual behavioural data provided by a videogame company from players in *Dead by Daylight* to test how hunting expertise and prey behaviour shape predator foraging specialization. *Dead by Daylight* is an online videogame simulating a predator-prey interaction, where one predator player hunts four prey players in different virtual environments. The data grants a high degree of precision on the behavioural interaction as both the predator and the four prey were measured simultaneously in each trial. First, we investigate how predators developped their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise because they encounter varying levels of difficulty with the prey that they hunt. Second, we test the hypothesis that individual variation in foraging behaviour will change with expertise. If expertise reduces the costs of switching between hunting tactics, we predict that individual predators should become more flexible with time. Alternatively, if expertise enables the refinement of the hunting tactics, then individuals may instead specialize. Third, we evaluate how predator foraging specialization interacts with variation in prey behaviour. Whether predators specialize or not with expertise may depend on the behaviour of their prey, for instance, as it may be harder to specialize when encounters are less predictable. Lastly, if individuals differ in their degree of foraging specialization, then we expect that the success of specialists and generalists will be equal. Specialist hunters should fare better when prey variability is lower, while flexible hunters should fare better when prey variability is higher.
